Trends of SDR in Thessaloniki

### 2023-06-19 Article_B145.Rmd
Commit 2023-06-19 14:17

```{r child = '/home/athan/CS_id/Clear_sky_id_Reno-Hansen_description_v14.2.Rmd'}
```

>>>>

<!-- CS_id v14.2 inserted here -->
```{r setup, echo=F, include=T, eval=TRUE}
## Load Clear Sky id variables
MS <- data.frame(readRDS("~/CS_id/PARAMS/Clear_sky_id_Reno-Hansen_apply_v14_2.Rds"))
# export all variables
for (an in names(MS)) {
    assign(an, MS[[an]])
}
```

<!-- # Detection of clear periods in GHI measurements for SDR trends. -->

As global radiation clear sky reference we are using the Haurwitz’s model, adjusted for our site with a factor of `r signif(alpha,3)` (Eq. \ref{eq:ahau}).
The selection of a clear sky reference model, was based on SDR observation from the period 2016 – 2021.
Where, after an iterative optimization of eight simple models (Daneshyar–Paltridge–Proctor, Kasten–Czeplak, Haurwitz, Berger–Duffie, Adnot–Bourges–Campana–Gicquel, Robledo-Soler, Kasten and Ineichen-Perez) with different factors.
We found, that Haurwitz’s model, adjusted with a factor of `r signif(alpha,3)` has the lower root mean squared error (RMSE). 
The tried models are described by @Reno2012 and tested by @Reno2016.
The iterative optimization method, for the selection of the reference is discussed by @Long2000 and @Reno2016.

\begin{equation}
\text{SDR}_\text{Clear Sky} = `r signif(alpha,3)` \times 1098 \times \cos( \text{SZA} ) \times \exp \left( \frac{ - 0.057}{\cos(\text{SZA})} \right) \label{eq:ahau}
\end{equation}

The following criteria and thresholds were used to identify clear-sky conditions.
Each criterion was applied for a running window of $`r nt`$ consecutive measurements/minutes, and the characterization is applied at the center value of the window.
A data point is consider as under cloud-sky condition if it fails to pass any of the criteria, all the other data points are characterized as clear-sky.

<!-- 1. MeanVIP -->
### Mean value of irradiance during the time period.

The mean of the measured value $\overline{G}_i$ must be inside an envelope based on the reference model $\text{SDR}_\text{Clear Sky}$ (Eq. \ref{eq:MeanVIP}).

\begin{equation}
`r 1 - CS_ref_rm_VIP_RelLow` \times \overline{\text{SDR}}_{i\text{Clear Sky}} - `r CS_ref_rm_VIP_LowOff`
< \overline{G}_i <
`r 1 + CS_ref_rm_VIP_RelUpp` \times \overline{\text{SDR}}_{i\text{Clear Sky}} + `r CS_ref_rm_VIP_UppOff`
\label{eq:MeanVIP}
\end{equation}

<!-- 2. MaxVIP -->
### Max value of irradiance during the time period.

The running max measured value $M_{Gi} = max[\text{SDR}_{i}]$, is compared to a similar constructed value from the reference $M_{CSi} = max[\text{SDR}_{i\text{Clear Sky}}]$ (Eq. \ref{eq:MaxVIP}).

\begin{equation}
`r MaxVIP_fct` \times M_{CSi} - `r MaxVIP_off_low`
< M_{Gi} <
`r MaxVIP_fct` \times M_{CSi} + `r MaxVIP_off_upp`
\label{eq:MaxVIP}
\end{equation}

<!-- 3. VIL -->
### Variability in irradiance by line length.

The length $L$ (Eq. \ref{eq:VILeq}) of the sequence of line segments connecting the points of the SDR time
series for the measured values $L$ and similar for the reference $L_{CS}$, must be within the limits of Eq. \ref{eq:VILcr}.

\begin{equation}
L = \sum_{i=1}^{n-1}\sqrt{\left ( \text{SDR}_{i+1} - \text{SDR}_{i}\right )^2 + \left ( t_{i+1} - t_i \right )^2}
\label{eq:VILeq}
\end{equation}

\begin{equation}
`r MinVIL_fct` \times L_{CSi} - `r offVIL_dwl` < L_i < `r MaxVIL_fct` \times L_{CSi} + `r offVIL_upl`
\label{eq:VILcr}
\end{equation}

<!-- 4. VCT -->
### Variance of Changes in the Time series.

<!-- Standard deviation of rate of change in irradiance. -->
We calculate the standard deviation $\sigma$ of the slope
($s$) between sequential points in the time series, normalized by the average SDR during the time interval.

\begin{gather}
s_i = \frac{\text{SDR}_{i+1} - \text{SDR}_{i}}{t_{i+1} - t_i}, \forall i \in \left \{ 1, 2, \ldots, n-1 \right \} \label{eq:VCT1} \\
\bar{s} = \frac{1}{n-1} \sum_{i=1}^{n-1} s_i \label{eq:VCT2} \\
\sigma_i = \frac {\sqrt{\frac{1}{n-1} \sum_{i=1}^{n-1} \left( s_i - \bar{s} \right)^2} } {\bar{G_i}} \label{eq:VCT3}
\end{gather}

\begin{equation}
\sigma_i < `r offVCT`
\label{eq:VCTcr}
\end{equation}

<!-- 5. VSM -->
### Variability in the Shape of the irradiance Measurements.

The maximum difference $X$ between the change in measured irradiance and the change in clear sky
irradiance over each measurement interval.

\begin{gather}
x_i = \text{SDR}_{i+1} - \text{SDR}_{i} \forall i \in \left \{ 1, 2, \ldots, n-1 \right \} \label{eq:VSM1} \\
x_{CS,i} = \text{SDR}_{CS,i+1} - \text{SDR}_{CS,i} \forall i \in \left \{ 1, 2, \ldots, n-1 \right \} \label{eq:VSM2} \\
X_i = \max{\left \{ \left | x_i - x_{CS,i} \right | \right \}} \label{eq:VSM3}
\end{gather}

\begin{equation}
X_i < `r offVSM`
\label{eq:VSMcr}
\end{equation}
